Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-09-04 papers/paper/ms.tex
update ms

Despite this and other observational challenges, young close-in planet
discovery has matured over the past decade, primarily due to Kepler,
K2, TESS, and Gaia
\citep[e.g.][]{Meibom_2013,Mann_K2_25_2016,Curtis_2018,Livingston_2018,Bouma_2020_toi837,Plavchan_2020,Newton_2021,Nardiello_2022,Barber_2022,Zhou_2022,Zakhozhay_2022,Wood_2023}.
The strategy pursued by most groups during the 2010s was to focus on
stars with known ages---obvious members of open clusters---and to
search them for transiting planets.  The resulting stellar, and
assumed planetary, ages are precise at the $\approx$10\% level.  
An alternative strategy, facilitated by Gaia, is to select transiting
planets based on empirical youth indicators, and to then search their local
neighborhoods for comoving companions \citep[e.g.][]{Tofflemire_2021}.

@@

Rotation-based ages have been reported for various subsets of Kepler
stars since the early data releases
\citep[e.g.][]{Walkowicz_2013,McQuillan_2014,Reinhold_2015,Angus_2018}.
More recent work has further explored incorporating information from
stellar kinematics \citep{2021AJ....161..189L,2024AJ....167..159L},
and from stellar colors, luminosities, and starspot amplitudes
\citep{2023ApJ...952..131M}.  Our analysis is motivated by a few
factors that can yield improvements, particularly for sub-gigayear
stars.   These factors are as follows.

@@

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.42\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
  \caption{{\bf The stars.}  Our analysis focuses on stars observed by
  Kepler with previously reported rotation periods (gray points).  The
  rotation periods are primarily drawn from
  \citet{Santos_2019,Santos_2021}.  About half of the stars with
  rotation periods are suitable for gyrochronology (dark gray points),
  based on factors including their non-binarity and proximity to the
  main sequence (orange line in \replaced{lower}{upper} panel; see
  Section~\ref{subsec:flags}).  Some host ``confirmed'' or
  ``candidate'' KOIs that meet additional planetary quality criteria
  (yellow points; Section~\ref{subsec:plflags}).  Surface gravities
  and effective temperatures were derived photometrically by
  \citet{Berger_2020a_catalog}.  Isochrones in the \replaced{lower}{upper} panel are
  from MIST \citep{Choi_2016}.
	}
	\label{fig:stellarprops}
\end{figure}

@@

To select stars with rotation periods, we turn to previous work.  Many
investigators have derived rotation periods both specifically for KOIs
\citep{McQuillan_2013,Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021},
and also for the broader set of all Kepler target stars
\citep{McQuillan_2014,Reinhold_2015,Santos_2019,Santos_2021,Reinhold2023}.
These studies used a range of detection methods and selection
functions.  We are interested in understanding the age distribution of
all Kepler target stars irrespective of KOI status.  Out of these
studies, the most homogeneous \deleted{and precise }analyses of both Kepler
targets and KOIs appear to be those by
\citet{Santos_2019,Santos_2021}, hereafter \citetalias{Santos_2019}
and \citetalias{Santos_2021}.  For discussion of the work by
\citet{Reinhold2023}, see Appendix~\ref{app:reinhold23}.
\citetalias{Santos_2019} and \citetalias{Santos_2021} combined a
wavelet analysis and autocorrelation-based approach, and cumulatively
reported rotation periods for 55{,}232 main-sequence and subgiant FGKM
stars.  They included known KOIs and binaries in their analysis, and
removed transits and eclipses during the stellar rotation measurement
process.

@@

{\it Near the main sequence (bit 9)}---Figure~\ref{fig:stellarprops} shows
that many stars observed by Kepler
are far from the main sequence.  Some of the challenges this introduces
for rotation-based ages include unresolved binaries, metallicity,
reddening, and drivers of rotation other than magnetic braking.
After exploring various options, we settled on the orange locus in the
$\log g$--$T_{\rm eff}$ plane shown in Figure~\ref{fig:stellarprops}
as a way of flagging unresolved binaries, as well as evolved
late-F and early-G stars.  While the exact details of how this locus
is constructed are arbitrary (see Appendix~\ref{app:linemethod}), the
general aim is to flag stars for which there is evidence based on
their location in the \replaced{HR}{HR or Kiel} diagram\added{s} that our gyrochronology model may not
be reliable.  While the $\log g$--$T_{\rm eff}$ cut flags about half
of Kepler stars with rotation periods as potentially not being
suitable for gyrochronology, a few outliers in $M_{\rm G}$ vs.~$G_{\rm
BP}$-$G_{\rm RP}$ remained after applying it.  Such outliers are
likely also questionable.  We therefore also fitted a polynomial to
the KOI main sequence in $M_{\rm G}$ vs.~$G_{\rm
	BP}$-$G_{\rm RP}$, and flagged stars more than 1\,magnitude from
this locus as part of the same bitflag.

%The rotation period
%distribution of sources excluded in this manner includes a relatively
%larger fraction of ``rapid rotators'' than the main sequence sources,
%likely due to a combination of astrophysical and systematic effects.
%See for instance Figure~8a of \citet{2022AJ....164..137K} or
%Figures~12 and 13 of \citet{2023ApJS..268....4F} for examples of how
%the period distribution of this population differs from stars near the
%main sequence.
\added{{\it Metallicity (bit 10)}---Theoretical models
	suggest that stars with non-solar metallicities
	spin down at different rates than solar-metallicity stars 
	\citep{Amard2020}.   This expectation has been hard
	to verify due to a dearth of nearby non-solar metallicity open clusters.
	However, if true, this effect could cause systematic
	biases for gyrochrone ages \citep[e.g.~Figure~9 of][]{Claytor2020}.
	We therefore flagged stars with spectroscopic $|[{\rm Fe/H}]|$$>$0.3
	from either LAMOST DR5 \citep{Zong2018} or CKS \citep{Petigura_2022}
	as being outside the range of our gyrochronology calibration.
	We discuss limitations of this approach in Section~\ref{sec:disc}.
}

{\it Candidate pulsators and close-in binaries (bit
\replaced{10}{11})}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g.\ Cepheids) and close-in binaries.
Visual inspection shows that although this flag selects many bona fide
objects in these classes, it also selects most known young planets
with $P_{\rm rot}$$\lesssim$5\,day.  These objects are
neither classical pulsators nor close binaries.  While we propagated
this flag into our set of quality flags, we therefore generally ignore
it.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.95\textwidth]{olympic_merged_lit_ages_vs_cluster_ages.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Reported ages of individual stars in benchmark open clusters.}
    Each point denotes a star's reported age, normalized by the age of
    a stellar ensemble in which the star is a candidate member.
    Cluster membership was evaluated without knowledge of rotation\added{, and for this plot,
    no attempt was made to exclude known binaries}.
    Although some field interlopers may be present in the membership
    lists, outliers can be compared between different methods on a
    relative basis.  Each study was cross-matched against the same
    cluster list; certain methods report ages for more stars than
    others.  Horizontal scatter is added to visually clarify the
    statistical age uncertainties.  At $<$1\,Gyr, our ages are
    generally accurate, and their statistical uncertainties match
    their dispersion.
	}
	\label{fig:agescalecompone}
\end{figure*}

@@

In Figure~\ref{fig:agescalecompone}, we compare our rotation-based
ages, and available ages from the literature, against the ages of
stars in these open clusters. 
From the literature, we
drew ages from \citet{Reinhold_2015}, \citet{Berger_2020a_catalog},
\citet{2021AJ....161..189L}, \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.  We followed any guidance available from
each study for removing ages that were unreliable, and plotted stars
within $1\sigma$ of zero as upper limits.  For the
\citeauthor{Reinhold_2015} ages, we used those calculated using the
\citet{Mamajek_2008} calibration; for the
\citeauthor{2023ApJ...952..131M} ages, we show those from
\texttt{kiauhoku} \citep{Claytor2020} rather than \texttt{STAREVOL}
\citep{Amard2019}, since the former showed better agreement with the
cluster age scale.

@@

Figure~\ref{fig:agescalecompone} show that although previous studies
reported ages that agree with the cluster scale for $\gtrsim$1\,Gyr
stars, sub-gigayear stars have historically had their ages
overestimated by 0.3--2\,dex, with severely understated uncertainties.
For isochrone and kinematic ages, this is because these methods rely
on parameters that do not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotation-based ages in \citet{2023ApJ...952..131M} and
\citet{2024AJ....167..159L}, the discrepancy is caused by details of
the calibration methodologies.  The \texttt{kiauhoku} spin-down model
for instance has known ``stretches'' and ``compressions'' in its age
scale relative to observed open clusters \citep[see][Sec.
7.3]{2023ApJ...952..131M}.  The \citet{Mamajek_2008} calibration used
by \citet{Reinhold_2015} appears more accurate than the former two
models, because it was fitted to reproduce the open cluster sequences
known at the time.  However, the uncertainties from all of these
methods appear to be underestimated.  This is probably because they do
not marginalize over the range of rotation periods accessible to
sub-gigayear stars of a fixed mass and age.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{have more precise ages}.  The statistical uncertainty for a
  single average star at 1, 2, and 3\,Gyr is shown, and is identical across each
  row.  
  Finally, the right panel compares the rotation-derived ages
  against star formation histories derived using CMD fitting (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

@@

The similarity of the KOI host and Kepler star age distributions adds
some nuance to the argument that young transiting planets are hard to
detect due to the photometric variability of their host stars.  If
true, this statement \replaced{seems to}{might} hold only for the very youngest
($\lesssim$0.4\,Gyr) stars, where there is a marginal deficit in the
KOI host age distribution relative to the parent stellar sample.
Examining scatter plots of the planet properties as a function of
age, we similarly find that fewer systems at $t_{\rm
gyro}$$<$0.4\,Gyr are detected with $P_{\rm orb}$$\gtrsim$30\,days
than at $t_{\rm gyro}$$>$0.4\,Gyr.

We can quantify the relative counts of stars as a function of age by
labelling stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr as ``young'',
``intermediate-age'', and ``old''.  A simple counting exercise from
Figure~\ref{fig:hist_tgyro} tells us that there are
\ratioobtoybstars\ times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.  Focusing only on the tails of the
distributions (0-0.3\,Gyr and 2.7-3\,Gyr), the implication is that the
formation rate of stars in the Kepler field decreased by a factor of
\ratiosfr$\pm$\uncratiosfr\ over the past three billion years.  If
this decrease continues linearly into the future, then a simple
extrapolation would imply that in $\approx$1.4$\pm$0.3\,Gyr the rate
of new stars forming in the Galaxy will reach zero.

@@

Four stellar ensembles in the Kepler field have to date yielded a
total of fourteen transiting planets.  Our analysis blindly recovers
the youth of all of these planets.

{\it Cep-Her}---The Cep-Her complex ($\approx$40\,Myr) contains
Kepler-1627Ab, Kepler-1643b, Kepler-1974b and Kepler-1975\,Ab
\citep{Bouma_2022a,Bouma_2022b}.  Different sub-groups of the complex
vary in age by $\approx$50\% \citep{Kerr2024}.  The planets
themselves are all on close-in orbits (5-25\,days), with sizes from
2-4\,$R_\oplus$.  Our rotation and lithium ages agree with the cluster
age for Kepler-1627A, Kepler-1974, and Kepler-1975A.  For Kepler-1643,
the rotation and cluster ages agree, and the lithium age is
2.0$\sigma$ above the cluster age reported by \citet{Bouma_2022b} for
RSG-5.

{\it MELANGE-3}---\citet{Barber_2022} reported a $105\pm10$\,Myr
association in the Kepler field containing two transiting
mini-Neptunes: Kepler-970 and Kepler-1928.  We find rotation-based
ages of $t_{\rm gyro}=176_{-40}^{+120}$\,Myr and
$144_{-88}^{+104}$\,Myr for Kepler-970 and Kepler-1928 respectively.
The former rotation-based age being slightly older than the suggested
Pleiades age for the association agrees with Figure~4 of
\citet{Barber_2022}.

{\it Theia~520}---We find rotation-based ages of $\approx$350\,Myr
for Kepler-968 and Kepler-52.  Based on spatial and kinematic
clustering, these stars are candidate 
members of Theia~520 \citep{2019AJ....158..122K}.  The core of this cluster is also known
as UBC-1 \citep{2018A&A...618A..59C}.  Based on isochrone and
rotation-based age-dating, Theia~520 seems to be
$\approx$230$\pm$70\,Myr old
\citep{2024A&A...681A..13F}, which is within
1$\sigma$ of our $t_{\rm gyro}$ measurements.  Ongoing work by
\citet{Curtis2024} broadly finds that the two planet-hosting stars are
indeed members of this diffuse population.  Given the cluster-level
and individual-star evidence, this makes Kepler-968 and Kepler-52 the
youngest multiplanet systems currently known from the main Kepler
mission.

{\it NGC~6811}---\citet{Meibom_2013} reported the discovery of
Kepler-66 and -67b, two mini-Neptunes in NGC~6811 ($\approx$1\,Gyr).
We derive rotation-based ages for these systems of
\kepsixsixtgyro\,Myr and \kepsixseventgyro\,Myr for Kepler-66 and
-67b, respectively.  Kepler-66 ($T_{\rm eff}$$\approx$5900\,K, $P_{\rm
rot}$$\approx$10.4\,days) has a lower precision and a more asymmetric
posterior because of its slow spin-down rate.  Kepler-67 ($T_{\rm
eff}$$\approx$5100\,K, $P_{\rm rot}$$\approx$10.3\,days) is marginally
hotter than the later K dwarfs that are in ``stalled'' spin-down at
this time, enabling its rotation period to be diagnostic of its age.

@@

Figure~\ref{fig:rp_period_age_results} shows the planets and planet
candidates with $t_{\rm gyro}$$<$1\,Gyr at 2$\sigma$.  \replaced{This}{The stars in this} figure
include\deleted{s} \nconfirmedplyounggyrotwosigmanograzingnoruwe\ confirmed
Kepler planets, and \ncandidateplyounggyrotwosigmanograzingnoruwe\
candidate planets, all of which are listed in Table~\ref{tab:planets}.
These include \nearthshighq\ objects Earth-sized or smaller,
\njupitershighq\ Jovian-sized planet candidates, \nsubsaturnshighq\
super-Neptunes (4--10\,$R_\oplus$), \nminineptuneshighq\ mini-Neptunes
above the radius valley as parametrized by
\citealt{2018MNRAS.479.4786V}, and \nsuperearthshighq\ super-Earth
sized planets below the radius valley.  While most of these planets
have orbital periods below 50\,days, \nlongperiodhighq\ are on more
distant orbits.

Among these systems, four highlights are Kepler-1529, Kepler-1565,
Kepler-1312, and Kepler-1629.  Kepler-1529\,b is a
$\approx$100$\pm$50\,Myr mini-Neptune with well-measured rotation and
lithium ages.  Kepler-1565\,b is an analogous $\approx$170-230\,Myr
super-Earth.  Kepler-1312 is a $t_{\rm gyro}$=\kepthirteentwelve\,Myr
near-Solar analog with an Earth-sized planet on a one-day orbit, and a
mini-Neptune on a five-day orbit.  The Earth-sized planet,
Kepler-1312\,c, along with the near-identical Kepler-1561\,b ($t_{\rm
gyro}$=\kepfifteensixone\,Myr), rank among the youngest Earth-sized
planets currently known, comparable to planets such as HD~63433\,d
\citep[1.1\,$R_\oplus$, $414$$\pm$$23$\,Myr;][]{2024AJ....167...54C}
and TOI-1807\,b \citep[1.3\,$R_\oplus$,
180$\pm$40\,Myr;][]{2021AJ....162...54H}.  Kepler-1629\,b, with a size
just two-thirds that of Earth (0.67$\pm$0.05\,$R_\oplus$), is
marginally older, also based on rotation (\kepsixteentwonine\,Myr).
The lithium age from a reconnaissance TRES spectrum confirms the
point.

>>>>

Despite this and other observational challenges, young close-in planet
discovery has matured over the past decade, primarily due to Kepler,
K2, TESS, and Gaia
\citep[e.g.][]{Meibom_2013,Mann_K2_25_2016,Curtis_2018,Livingston_2018,Bouma_2020_toi837,Plavchan_2020,Newton_2021,Nardiello_2022,Barber_2022,Zhou_2022,Zakhozhay_2022,Wood_2023}.
The strategy pursued by most groups during the 2010s was to focus on
stars with known ages---obvious members of open clusters---and to
search them for transiting planets.  The resulting stellar, and
assumed planetary, ages are precise at the $\approx$10\% level.  
An alternative strategy, facilitated by Gaia, is to select \deleted{transiting}
planets based on empirical youth indicators, and to then search their local
neighborhoods for comoving \added{stellar }companions \citep[e.g.][]{Tofflemire_2021}.

@@

Rotational ages have been reported for various subsets of Kepler
stars since the early data releases
\citep[e.g.][]{Walkowicz_2013,McQuillan_2014,Reinhold_2015,Angus_2018}.
More recent work has further explored incorporating information from
stellar kinematics \citep{2021AJ....161..189L,2024AJ....167..159L},
and from stellar colors, luminosities, and starspot amplitudes
\citep{2023ApJ...952..131M}.  Our analysis is motivated by a few
factors that can yield improvements, particularly for sub-gigayear
stars.   These factors are as follows.

@@

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.42\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
  \caption{{\bf The stars.}  Our analysis focuses on stars observed by
  Kepler with previously reported rotation periods (gray points).  The
  rotation periods are primarily drawn from
  \citet{Santos_2019,Santos_2021}.  About half of the stars with
  rotation periods are suitable for gyrochronology (dark gray\deleted{ points}),
  based on factors including their \added{apparent }non-binarity and proximity to the
  main sequence (orange line in \replaced{lower}{upper} panel; see
  Section~\ref{subsec:flags}).  Some host ``confirmed'' or
  ``candidate'' KOIs that meet additional planetary quality criteria
  (yellow points; Section~\ref{subsec:plflags}).  Surface gravities
  and effective temperatures were derived photometrically by
  \citet{Berger_2020a_catalog}.  Isochrones in the \replaced{lower}{upper} panel are
  from MIST \citep{Choi_2016}.
	}
	\label{fig:stellarprops}
\end{figure}

@@

To select stars with rotation periods, we turn to previous work.  Many
investigators have derived rotation periods both specifically for KOIs
\citep{McQuillan_2013,Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021},
and also for the broader \replaced{set of all}{sample of} Kepler target stars
\citep{McQuillan_2014,Reinhold_2015,Santos_2019,Santos_2021,Reinhold2023}.
These studies used a range of detection methods and selection
functions.  We are interested in understanding the age distribution of
all Kepler target stars irrespective of KOI status.  Out of these
studies, the most homogeneous \deleted{and precise }analyses of both Kepler
targets and KOIs appear to be those by
\citet{Santos_2019,Santos_2021}, hereafter \citetalias{Santos_2019}
and \citetalias{Santos_2021}.  For discussion of the work by
\citet{Reinhold2023}, see Appendix~\ref{app:reinhold23}.
\citetalias{Santos_2019} and \citetalias{Santos_2021} combined a
wavelet analysis and autocorrelation-based approach, and cumulatively
reported rotation periods for 55{,}232 main-sequence and subgiant FGKM
stars.  They included known KOIs and binaries in their analysis, and
removed transits and eclipses during the stellar rotation measurement
process.

@@

{\it Near the main sequence (bit 9)}---Figure~\ref{fig:stellarprops} shows
that many stars observed by Kepler
are far from the main sequence.  Some of the challenges this introduces
for rotational ages include unresolved binaries, metallicity,
reddening, and drivers of rotation other than magnetic braking.
After exploring various options, we settled on the orange locus in the
$\log g$--$T_{\rm eff}$ plane shown in Figure~\ref{fig:stellarprops}
as a way of flagging unresolved binaries, as well as evolved
late-F and early-G stars.  While the exact details of how this locus
is constructed are arbitrary (see Appendix~\ref{app:linemethod}), the
general aim is to flag stars for which there is evidence based on
their location in the \replaced{HR}{HR or Kiel} diagram\added{s} that our gyrochronology model may not
be reliable.  While the $\log g$--$T_{\rm eff}$ cut flags about half
of Kepler stars with rotation periods as potentially not being
suitable for gyrochronology, a few outliers in $M_{\rm G}$ vs.~$G_{\rm
BP}$-$G_{\rm RP}$ remained after applying it.  Such outliers are
likely also questionable.  We therefore also fitted a polynomial to
the KOI main sequence in $M_{\rm G}$ vs.~$G_{\rm
	BP}$-$G_{\rm RP}$, and flagged stars more than 1\,magnitude from
this locus as part of the same bitflag.

%The rotation period
%distribution of sources excluded in this manner includes a relatively
%larger fraction of ``rapid rotators'' than the main sequence sources,
%likely due to a combination of astrophysical and systematic effects.
%See for instance Figure~8a of \citet{2022AJ....164..137K} or
%Figures~12 and 13 of \citet{2023ApJS..268....4F} for examples of how
%the period distribution of this population differs from stars near the
%main sequence.
\added{{\it Metallicity (bit 10)}---Models
	suggest that stars with non-solar metallicities
	spin down at different rates than solar-metallicity stars 
	\citep{Amard2020}.   This expectation has been hard
	to verify due to a dearth of nearby non-solar metallicity open clusters.
	However, if true, this could cause systematic
	biases for gyrochrone ages \citep[e.g.~Figure~9 of][]{Claytor2020}.
	We therefore flagged stars with spectroscopic $|[{\rm Fe/H}]|$$>$0.3
	from either LAMOST DR5 \citep{Zong2018} or CKS \citep{Petigura_2022}
	as being outside the range of our gyrochronology calibration.
	We discuss limitations of this approach in Section~\ref{sec:disc}.
}

{\it Candidate pulsators and close-in binaries (bit
\replaced{10}{11})}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g.\ Cepheids) and close-in binaries.
Visual inspection shows that although this flag selects many bona fide
objects in these classes, it also selects known young planets
with $P_{\rm rot}$$\lesssim$5\,day.  These objects are
neither classical pulsators nor close binaries.  While we propagated
this flag into our set of quality flags, we therefore generally ignore
it.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.95\textwidth]{olympic_merged_lit_ages_vs_cluster_ages.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Reported ages of individual stars in benchmark open clusters.}
    Each point denotes a star's reported age, normalized by the age of
    a stellar ensemble in which the star is a candidate member.
    Cluster membership was evaluated without knowledge of rotation\added{, and for this plot,
    no attempt was made to exclude known binaries}.
    Although some field interlopers may be present in the membership
    lists, outliers can be compared between different methods on a
    relative basis.  Each study was cross-matched against the same
    cluster list; certain methods report ages for more stars than
    others.  Horizontal scatter is added to visually clarify the
    statistical age uncertainties.  Below 1\,Gyr, our ages are
    generally accurate, and their statistical uncertainties match
    their dispersion.
	}
	\label{fig:agescalecompone}
\end{figure*}

@@

In Figure~\ref{fig:agescalecompone}, we compare our rotational
ages, and available ages from the literature, against the ages of
stars in these open clusters. 
From the literature, we
drew ages from \citet{Reinhold_2015}, \citet{Berger_2020a_catalog},
\citet{2021AJ....161..189L}, \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.  We followed any guidance available from
each study for removing ages that were unreliable, and plotted stars
within $1\sigma$ of zero as upper limits.  For the
\citeauthor{Reinhold_2015} ages, we used those calculated using the
\citet{Mamajek_2008} calibration; for the
\citeauthor{2023ApJ...952..131M} ages, we show those from
\texttt{kiauhoku} \citep{Claytor2020} rather than \texttt{STAREVOL}
\citep{Amard2019}, since the former showed better agreement with the
cluster age scale.

@@

Figure~\ref{fig:agescalecompone} show that although previous studies
reported ages that agree with the cluster scale for $\gtrsim$1\,Gyr
stars, sub-gigayear stars have historically had their ages
overestimated by 0.3--2\,dex, with severely understated uncertainties.
For isochrone and kinematic ages, this is because these methods rely
on parameters that do not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotational ages in \citet{2023ApJ...952..131M} and
\citet{2024AJ....167..159L}, the discrepancy is caused by details of
the calibration methodologies.  The \texttt{kiauhoku} spin-down model
for instance has known ``stretches'' and ``compressions'' in its age
scale relative to observed open clusters \citep[see][Sec.
7.3]{2023ApJ...952..131M}.  The \citet{Mamajek_2008} calibration used
by \citet{Reinhold_2015} appears more accurate than the former two
models, because it was fitted to reproduce the open cluster sequences
known at the time.  However, the uncertainties from all of these
methods appear to be underestimated.  This is probably because they do
not marginalize over the range of rotation periods accessible to
sub-gigayear stars of a fixed mass and age.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{have more precise ages}.  The
  statistical uncertainty for an average star at 1, 2, and 3\,Gyr is
  shown, and is identical across each
  row.  
  Finally, the right panel compares the rotation-derived ages
  against star formation histories derived using CMD fitting (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

@@

The similarity of the KOI host and Kepler star age distributions adds
some nuance to the argument that young transiting planets are hard to
detect due to the photometric variability of their host stars.  If
true, this statement \replaced{seems to}{might} hold only for the very youngest
($\lesssim$0.4\,Gyr) stars, where there is a marginal deficit in the
KOI host age distribution relative to the parent stellar sample.
Examining scatter plots of the planet properties as a function of
age, we similarly find that fewer systems at $t_{\rm
gyro}$$<$0.4\,Gyr are detected with orbital periods beyond 30\,days
than at $t_{\rm gyro}$$>$0.4\,Gyr.

We can quantify the relative counts of stars as a function of age by
labelling stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr as ``young'',
``intermediate-age'', and ``old''.  A simple counting exercise from
Figure~\ref{fig:hist_tgyro} tells us that there are
\ratioobtoybstars\ times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.  Focusing only on the tails of the
distributions (0-0.3\,Gyr and 2.7-3\,Gyr), the implication is that the
formation rate of stars in the Kepler field decreased by a factor of
\ratiosfr$\pm$\uncratiosfr\ over the past three billion years.  If
this decrease continues linearly into the future, then a simple
extrapolation would imply that in $\approx$1.4$\pm$0.3\,Gyr the
abundance of newly-formed stars in this region of the Galaxy will
reach zero.

@@

Four stellar ensembles in the Kepler field have to date yielded
\deleted{a total of }fourteen transiting planets.  Our analysis
blindly recovers the youth of all of these planets.

{\it Cep-Her}---The Cep-Her complex ($\approx$40\,Myr) contains
Kepler-1627Ab, Kepler-1643b, Kepler-1974b and Kepler-1975\,Ab
\citep{Bouma_2022a,Bouma_2022b}.  Different sub-groups of the complex
vary in age by $\approx$50\% \citep{Kerr2024}.  The planets
themselves are all on close-in orbits (5-25\,days), with sizes from
2-4\,$R_\oplus$.  Our rotation and lithium ages agree with the cluster
age for Kepler-1627A, Kepler-1974, and Kepler-1975A.  For Kepler-1643,
the rotation and cluster ages agree, and the lithium age is
2.0$\sigma$ above the cluster age \replaced{reported by}{from} \citet{Bouma_2022b} for
RSG-5.

{\it MELANGE-3}---\citet{Barber_2022} reported a $105$$\pm$$10$\,Myr
association in the Kepler field containing two transiting
mini-Neptunes: Kepler-970 and Kepler-1928.  We find rotational
ages of $t_{\rm gyro}=176_{-40}^{+120}$\,Myr and
$144_{-88}^{+104}$\,Myr for Kepler-970 and Kepler-1928 respectively.
The former \deleted{rotation-based }age being slightly older than the suggested
Pleiades age for the association agrees with Figure~4 of
\citet{Barber_2022}.

{\it Theia~520}---We find rotational ages of $\approx$350\,Myr
for Kepler-968 and Kepler-52.  Based on spatial and kinematic
clustering, these stars are candidate 
members of Theia~520 \citep{2019AJ....158..122K}.  The core of this cluster is also known
as UBC-1 \citep{2018A&A...618A..59C}.  Based on isochrone and
rotational age-dating, Theia~520 seems to be
$\approx$230$\pm$70\,Myr old
\citep{2024A&A...681A..13F}, which is within
1$\sigma$ of our $t_{\rm gyro}$ measurements.  Ongoing work by
\citet{Curtis2024} broadly finds that the two planet-hosting stars are
indeed members of this diffuse population.  Given the cluster-level
and individual-star evidence, this makes Kepler-968 and Kepler-52 the
youngest multiplanet systems currently known from the main Kepler
mission.

{\it NGC~6811}---\citet{Meibom_2013} reported the discovery of
Kepler-66 and -67b, two mini-Neptunes in NGC~6811 ($\approx$1\,Gyr).
We derive rotational ages for these systems of
\kepsixsixtgyro\,Myr and \kepsixseventgyro\,Myr for Kepler-66 and
-67b, respectively.  Kepler-66 ($T_{\rm eff}$$\approx$5900\,K, $P_{\rm
rot}$$\approx$10.4\,days) has a lower precision and a more asymmetric
posterior because of its slow spin-down rate.  Kepler-67 ($T_{\rm
eff}$$\approx$5100\,K, $P_{\rm rot}$$\approx$10.3\,days) is marginally
hotter than the later K dwarfs that are in ``stalled'' spin-down at
this time, enabling its rotation period to be diagnostic of its age.

@@

Figure~\ref{fig:rp_period_age_results} shows the planets and planet
candidates with $t_{\rm gyro}$$<$1\,Gyr at 2$\sigma$.  \replaced{This}{The stars in this} figure
include\deleted{s} \nconfirmedplyounggyrotwosigmanograzingnoruwe\ confirmed
Kepler planets, and \ncandidateplyounggyrotwosigmanograzingnoruwe\
candidate planets, all \deleted{of which are }listed in Table~\ref{tab:planets}.
These include \nearthshighq\ objects Earth-sized or smaller,
\njupitershighq\ Jovian-sized planet candidates, \nsubsaturnshighq\
super-Neptunes (4--10\,$R_\oplus$), \nminineptuneshighq\ mini-Neptunes
above the radius valley as parametrized by
\citealt{2018MNRAS.479.4786V}, and \nsuperearthshighq\ super-Earth
sized planets below the radius valley.  While most of these planets
have orbital periods below 50\,days, \nlongperiodhighq\ are on more
distant orbits.

Among these systems, four highlights are Kepler-1529, Kepler-1565,
Kepler-1312, and Kepler-1629.  Kepler-1529\,b is a
$\approx$100$\pm$50\,Myr mini-Neptune with well-measured rotation and
lithium ages.  Kepler-1565\,b is an analogous $\approx$170-230\,Myr
super-Earth.  Kepler-1312 is a $t_{\rm gyro}$=\kepthirteentwelve\,Myr
near-Solar analog with an Earth-sized planet on a one-day orbit, and a
mini-Neptune on a five-day orbit.  The Earth-sized planet,
Kepler-1312\,c, along with the near-identical Kepler-1561\,b ($t_{\rm
gyro}$=\kepfifteensixone\,Myr), \replaced{rank among}{could be} the youngest Earth-sized
planets currently known, \added{and are }comparable to planets such as HD~63433\,d
\citep[1.1\,$R_\oplus$, $414$$\pm$$23$\,Myr;][]{2024AJ....167...54C}
and TOI-1807\,b \citep[1.3\,$R_\oplus$,
180$\pm$40\,Myr;][]{2021AJ....162...54H}.  Kepler-1629\,b, with a size
just two-thirds that of Earth (0.67$\pm$0.05\,$R_\oplus$), is
marginally older, also based on rotation (\kepsixteentwonine\,Myr).
The lithium age from a reconnaissance TRES spectrum confirms the
point.
